---
title: Data wrangling - The New Sound of Politics
authors:
  - name: Robert Bowien
    affiliation: University College Dublin
    roles: writing
    corresponding: true
bibliography: references.bib
---

This document shows the data wrangling for the Project "The New Sound of Politics".

First, the relevant libraries and data are loaded. Then, time-series data from the GLES is identified and isolated.

```{r}
library(ggplot2)
library(dplyr)
library(stringfish)
library(stringr)
```


# Part 1: Social media 
```{r}

# gles_data <- haven::read_dta("data/ZA6832_v2-0-0.dta")
# gles_korresp <- readxl::read_xlsx("data/ZA6832_v2-0-0_Korrespondenzliste.xlsx")

```

```{r}
# nrow(gles_data)
# nrow(gles_korresp)

# print(gles_korresp[grepl("Pol. Mediennutzun", gles_korresp$Variablenlabel),c("Variablenname", "Variablenlabel")])

# colnames(gles_data)

# info_source_data <- gles_data |> subset(select = c("field_start", "field_end", "sample", "version", "lfdn", "lfdn_od", "m0001_1", "m0001_2", "m0001_3")) # m0001_1 does not yet include social media, social media in there since 2012



```
Now, we have a first look at the data. Mainly interesting to us is the data that included social media as a seperate category. We want to know:
- from when on is social media included?
- how is the data distributed for the oldest and the latest year?
- how did it develop over time?
```{r}
# print(info_source_data$m0001_1[1:10])
# print(info_source_data$m0001_2[1:10])
# print(info_source_data$m0001_3[1:10])

# info_source_data[info_source_data$m0001_3 %in% c(-99:-71), ] <- NA

# info_source_3 <- info_source_data[!is.na(info_source_data$m0001_3), ]

# write.csv(info_source_3, "data/sm_relevance.csv")

# From when on is social media included?

# min(info_source_3$field_start)
# unique(info_source_3$field_start)
# length(unique(info_source_3$lfdn_od))
# nrow(info_source_3)

# info_source_3[info_source_3$lfdn_od == 30, ]
# # --> the respondent id is lfdn_od

# data_hist_sample54 <- info_source_3[info_source_3$sample == 54, ]
# data_hist_sample46 <- info_source_3[info_source_3$sample == 46, ]

# hist(data_hist_sample46$m0001_3)
# hist(data_hist_sample54$m0001_3)


```


We can observe a clear positive trend of the relevance of social media as a political information source.


# Part 2: Bundestag speeches

Here, the speeches in the Bundestag are cleaned and filtered as good as possible. However, the structure and lack of tidyness of the data pose challenging problems, some of which could not be resolved to adhere to the highest scientific standards.

```{r}
bt_speeches_dat <- read.csv("data/Bundestag.csv")
bt_speeches <- bt_speeches_dat[bt_speeches_dat$Period >= 20,]

# bt_speeches_test <- bt_speeches |> sample_n(1000)
# write.csv(x=bt_speeches_test, "data/Bundestag_testset.csv")
# bt_speeches <- read.csv("data/bundestag_181920.csv")
bt_meta <- read.csv("data/all_mps_meta.csv", sep = ";")
```

```{r}

bt_speeches$speech_id <- seq_len(nrow(bt_speeches))
bt_speeches[, c("Session", "Date", "Interjection", "MPID", "Party", "speech_id")] |> head()

bt_speeches_meta <- merge(bt_speeches, bt_meta, by="MPID", all.x = TRUE) |> arrange(speech_id)
bt_speeches_meta[, c("Session", "Date", "Interjection", "MPID", "Party", "speech_id")] |> head()

bt_speeches_nochair <-  bt_speeches_meta[bt_speeches_meta$Chair == "False", ]

bt_speeches_nointerjections <- bt_speeches_nochair[bt_speeches_nochair$Interjection == "False", ]
bt_speeches_nointerjections[, c("Session", "Date", "Interjection", "MPID", "Party", "speech_id")] |> head()

```

```{r}

bt_speeches_nounidentified <- bt_speeches_nointerjections |> dplyr::filter(MPID != 0)
bt_speeches_nounidentified[, c("Session", "Date", "Chair", "Interjection", "MPID", "Party", "speech_id")] |> head()


bt_speeches_combined <- bt_speeches_nounidentified |> group_by(MPID, Date, Chair, Session, Period, Party, Name) |>
summarise(Speech = paste(Speech, collapse="")) # if wanted, insert pattern in collapse = "" to see where the splits occur afterwards. 

bt_speeches_combined |> nrow() |> print()
bt_speeches_combined$Date |> min() |> print()
names(bt_speeches_nounidentified)


# names_bt_members <- bt_meta$Name |> unique() |> paste0(collapse = "|")
names_bt_members_list <- unique(bt_meta$Name) |> list()
names_bt_members <- split(names_bt_members_list[[1]], ceiling(seq_along(names_bt_members_list[[1]])/300))


names_bt_members_rgx <- sapply(names_bt_members, function(names){
  gsub(" \\(.+\\)", "", paste0(names, collapse="|"))
})

```
```{r}

# Disclaimer: following regex written by author
remove_regex_base = "(\\d{4} )?Deutscher Bundestag – \\d{2} ?. Wahlperiode – \\d+ ?. Sitzung ?. Berlin, ((Mon|Diens|Donners|Frei|Sams|Sonn)tag|Mittwoch), den \\d+ ?. (Januar|Februar|M(ä|ae)rz|April|Mai|Juni|Juli|August|September|Oktober|November|Dezember) \\d{4,}( \\d*)?"

# Disclaimer: following regex written by claude
lookaround_regexes_rm_names_b4 = sapply(names_bt_members_rgx, function(regex) {paste0(paste0(
    "(?:Präsident\\s+Dr\\.\\s+)?",        # "Präsident Dr." is optional
    "(?:", regex, ")",              # any name from the pipe-separated list
    "\\s+Deutscher\\s+Bundestag",
    "\\s+[–-]\\s+\\d+\\s*\\.\\s*Wahlperiode",
    "\\s+[–-]\\s+\\d+\\s*\\.\\s*Sitzung",
    "\\s*\\.\\s*Berlin,",
    "\\s+\\w+,\\s+den\\s+\\d+\\s*\\.\\s*\\w+\\s+\\d{4}",
    "\\s+\\d+"
  ))})

# Disclaimer: following regex written by perplexity
lookaround_regexes_rm_names_bhind = sapply(names_bt_members_rgx, function(regex) {paste0("(?:", "(?:(?:\\d{2,4} )?Deutscher Bundestag – \\d{2} ?. Wahlperiode – \\d+ ?. Sitzung ?. Berlin, (?:(?:Mon|Diens|Donners|Frei|Sams|Sonn)tag|Mittwoch), den \\d+ ?. (?:Januar|Februar|M(?:ä|ae)rz|April|Mai|Juni|Juli|August|September|Oktober|November|Dezember) \\d{4,}(?: \\d*)?) (?:Alterspr(?:ä|ae)sident|Parl\\. Staatssekret(?:ä|ae)r)? ?(?:\\w{2,4}\\. ){0,2}) ?(Alterspr(ä|ae)sident|Parl. Staatssekret(ä|ae)r)? ?(\\w{2,4}\\. ){0,2}(", regex,")")})

bt_speeches_combined_clean <- bt_speeches_combined

```


```{r}

removeRegexes <- function(column, regex, n_threads = 4){
  for (n in seq_along(regex)){
    rgx <- regex[n]
    print(paste0("rgx group: ", n, " of ", length(regex)))
    # Source - https://stackoverflow.com/a/78270004
    # Posted by LMc
    # Retrieved 2026-04-08, License - CC BY-SA 4.0
    column <- sf_gsub(column, regex(rgx, ignore_case = TRUE), "", nthreads = n_threads)
  
  }
  return(column)
}

bt_speeches_combined_clean$Speech <- removeRegexes(bt_speeches_combined_clean$Speech, lookaround_regexes_rm_names_b4, 8)

bt_speeches_combined_clean$Speech <- removeRegexes(bt_speeches_combined_clean$Speech, lookaround_regexes_rm_names_bhind, 8)

bt_speeches_combined_clean$Speech <- removeRegexes(bt_speeches_combined_clean$Speech, remove_regex_base, 8)



```

```{r}
bt_speeches_combined_clean |> head()

```



```{r}
write.csv(bt_speeches_combined_clean, "data/bundestag_cleaned_period20.csv")

```
The dataset bundestag_seit2020 contains 
```{r}
22137
```

speeches from 2020-01-15 to 2023-04-26

The dataset bundestag_181920 contains

```{r}
nrow(bt_speeches_combined_clean)
min(bt_speeches_combined_clean$Date)
max(bt_speeches_combined_clean$Date)

```
speeches from 2013-10-22 to 2023-04-26


