---
title: Data wrangling - The New Sound of Politics
authors:
  - name: Robert Bowien
    affiliation: University College Dublin
    roles: writing
    corresponding: true
bibliography: references.bib
---

This document shows the data wrangling for the Project "The New Sound of Politics".

First, the relevant libraries and data are loaded. Then, time-series data from the GLES is identified and isolated.

```{r}
library(ggplot2)
library(dplyr)
```

```{r}

gles_data <- haven::read_dta("new-sound-of-politics/data/ZA6832_v2-0-0.dta")
gles_korresp <- readxl::read_xlsx("new-sound-of-politics/data/ZA6832_v2-0-0_Korrespondenzliste.xlsx")

```

```{r}
nrow(gles_data)
nrow(gles_korresp)

print(gles_korresp[grepl("Pol. Mediennutzun", gles_korresp$Variablenlabel),c("Variablenname", "Variablenlabel")])

colnames(gles_data)

info_source_data <- gles_data |> subset(select = c("field_start", "field_end", "sample", "version", "lfdn", "lfdn_od", "m0001_1", "m0001_2", "m0001_3")) # m0001_1 does not yet include social media, social media in there since 2012



```
Now, we have a first look at the data. Mainly interesting to us is the data that included social media as a seperate category. We want to know:
- from when on is social media included?
- how is the data distributed for the oldest and the latest year?
- how did it develop over time?
```{r}
print(info_source_data$m0001_1[1:10])
print(info_source_data$m0001_2[1:10])
print(info_source_data$m0001_3[1:10])

info_source_data[info_source_data$m0001_3 %in% c(-99:-71), ] <- NA

info_source_3 <- info_source_data[!is.na(info_source_data$m0001_3), ]

# From when on is social media included?

min(info_source_3$field_start)
unique(info_source_3$field_start)
length(unique(info_source_3$lfdn_od))
nrow(info_source_3)

info_source_3[info_source_3$lfdn_od == 30, ]
# --> the respondent id is lfdn_od

data_hist_sample54 <- info_source_3[info_source_3$sample == 54, ]
data_hist_sample46 <- info_source_3[info_source_3$sample == 46, ]

hist(data_hist_sample46$m0001_3)
hist(data_hist_sample54$m0001_3)


```

The next step is to calculate the proportion of social media as most important information source and then plot the development over time.


```{r}


data_ts_linechart <- info_source_3 |> group_by(field_start) |> summarise(prop_sm = mean(m0001_3 == 4))

ggplot(data = data_ts_linechart, aes(x=field_start, y=prop_sm, group = 1)) +
  geom_line()


```