---
title: data analysis - The New Sound of Politics
authors:
  - name: Robert Bowien
    affiliation: University College Dublin
    roles: writing
    corresponding: true
bibliography: references.bib
---



# Descriptive Data analysis

## Social Media relevance

The next step is to calculate the proportion of social media as most important information source and then plot the development over time.


```{r}


data_ts_linechart <- info_source_3 |> group_by(field_start) |> summarise(prop_sm = mean(m0001_3 == 4))

ggplot(data = data_ts_linechart, aes(x=field_start, y=prop_sm, group = 1)) +
  geom_line()


```


## Bundestag Speeches


load the data
```{r}

bt_speeches <- read.csv("data/bundesta_seit2020.csv")


```
calculate summary statisitics


```{r}

bt_speeches <- bt_speeches |> group_by(Party) |> mutate(n_party_members = length(unique(MPID)))

sum_stat = subset(bt_speeches, select = c("Party", "n_party_members")) |> unique()

print(sum_stat)
print(bt_speeches[bt_speeches$n_party_members == max(bt_speeches$n_party_members),]$Party |> unique())

```
Note: If populist sentiment analysis is too difficult, go for pos/neg sentiment analysis --> neg. sentiment does better on social media, success strategy for right wing, positive sentiment more for left wing parties


