---
title: Data wrangling - The New Sound of Politics
authors:
  - name: Robert Bowien
    affiliation: University College Dublin
    roles: writing
    corresponding: true
bibliography: references.bib
---

This document shows the data wrangling for the Project "The New Sound of Politics".

First, the relevant libraries and data are loaded. Then, time-series data from the GLES is identified and isolated.

```{r}
library(ggplot2)
library(dplyr)
library(stringfish)
library(stringr)
```


# Part 1: Social media 
```{r}

# gles_data <- haven::read_dta("data/ZA6832_v2-0-0.dta")
# gles_korresp <- readxl::read_xlsx("data/ZA6832_v2-0-0_Korrespondenzliste.xlsx")

```

```{r}
# nrow(gles_data)
# nrow(gles_korresp)

# print(gles_korresp[grepl("Pol. Mediennutzun", gles_korresp$Variablenlabel),c("Variablenname", "Variablenlabel")])

# colnames(gles_data)

# info_source_data <- gles_data |> subset(select = c("field_start", "field_end", "sample", "version", "lfdn", "lfdn_od", "m0001_1", "m0001_2", "m0001_3")) # m0001_1 does not yet include social media, social media in there since 2012



```
Now, we have a first look at the data. Mainly interesting to us is the data that included social media as a seperate category. We want to know:
- from when on is social media included?
- how is the data distributed for the oldest and the latest year?
- how did it develop over time?
```{r}
# print(info_source_data$m0001_1[1:10])
# print(info_source_data$m0001_2[1:10])
# print(info_source_data$m0001_3[1:10])

# info_source_data[info_source_data$m0001_3 %in% c(-99:-71), ] <- NA

# info_source_3 <- info_source_data[!is.na(info_source_data$m0001_3), ]

# write.csv(info_source_3, "data/sm_relevance.csv")

# From when on is social media included?

# min(info_source_3$field_start)
# unique(info_source_3$field_start)
# length(unique(info_source_3$lfdn_od))
# nrow(info_source_3)

# info_source_3[info_source_3$lfdn_od == 30, ]
# # --> the respondent id is lfdn_od

# data_hist_sample54 <- info_source_3[info_source_3$sample == 54, ]
# data_hist_sample46 <- info_source_3[info_source_3$sample == 46, ]

# hist(data_hist_sample46$m0001_3)
# hist(data_hist_sample54$m0001_3)


```




We can observe a clear positive trend of the relevance of social media as a political information source.


# Part 2: Bundestag speeches

```{r}
bt_speeches_dat <- read.csv("data/Bundestag.csv")
bt_speeches <- bt_speeches_dat[bt_speeches_dat$Period >= 18,]

# bt_speeches <- read.csv("data/bundestag_181920.csv")
bt_meta <- read.csv("data/all_mps_meta.csv", sep = ";")
```

```{r}

bt_speeches_nointerjections <- bt_speeches[bt_speeches$Interjection == "False", ]

```
Do I need to concatenate the speeches or should I leave them like this? Yes because I want one document to be one speech and not one random part of a speech.
Note: some speakers could not be identified and have the MPID 0. These speakers will be excluded at this time.
```{r}

bt_speeches_nounidentified <- bt_speeches_nointerjections |> dplyr::filter(MPID != 0)

bt_speeches_combined <- bt_speeches_nounidentified |> group_by(MPID, Date, Chair, Session, Period, Party) |>
arrange(.bygroup=TRUE) |>
summarise(Speech = paste(Speech, collapse="-|-"))

bt_speeches_combined |> nrow() |> print()
bt_speeches_combined$Date |> min() |> print()
head(bt_speeches_combined)


# names_bt_members <- bt_meta$Name |> unique() |> paste0(collapse = "|")
names_bt_members_list <- unique(bt_meta$Name) |> list()
names_bt_members <- split(names_bt_members_list[[1]], ceiling(seq_along(names_bt_members_list[[1]])/300))

# chunk_size <- round(nchar(names_bt_members) / 20)
# names_bt_members_list <- substring(names_bt_members,
#                    first = seq(1, nchar(names_bt_members), by = chunk_size),
#                    last  = seq(chunk_size, nchar(names_bt_members), by = chunk_size))

names_bt_members_rgx <- sapply(names_bt_members, function(names){
  gsub(" \(.+\)", "", paste0(names, collapse="|"))
})



```
```{r}
remove_regex_base = "(\d{4} )?Deutscher Bundestag – \d{2} ?. Wahlperiode – \d+ ?. Sitzung ?. Berlin, ((Mon|Diens|Donners|Frei|Sams|Sonn)tag|Mittwoch), den \d+ ?. (Januar|Februar|M(ä|ae)rz|April|Mai|Juni|Juli|August|September|Oktober|November|Dezember) \d{4,}( \d*)?"


remove_regexes = sapply(names_bt_members_rgx, function(regex) {paste0("(", regex, ")? ?", remove_regex_base, "( (Alterspr(ä|ae)sident|Parl. Staatssekret(ä|ae)r) (\w{2,4}\. ){0,2}(", regex,")?)?")})

bt_speeches_combined_clean <- bt_speeches_combined

for (n in seq(1:length(remove_regexes))){
  rgx <- remove_regexes[n]
  print(paste0("rgx group: ", n, " of ", length(remove_regexes)))
  # Source - https://stackoverflow.com/a/78270004
  # Posted by LMc
  # Retrieved 2026-04-08, License - CC BY-SA 4.0
  bt_speeches_combined_clean$Speech <- sf_gsub(bt_speeches_combined_clean$Speech, rgx, "", nthreads = 4)
}


# remove_regex = paste0(remove_regex_base, remove_regex_add2)
# bt_speeches_combined_clean <- bt_speeches_combined_clean |> mutate(Speech = gsub(remove_regex, "", Speech))


```

```{r}


bt_speeches_combined_clean$speech_id <- seq_len(nrow(bt_speeches_combined_clean))

bt_speeches_meta <- merge(bt_speeches_combined_clean, bt_meta, by="MPID", all.x = TRUE) |> arrange(speech_id)


bt_speeches_meta[bt_speeches_meta$Name == "Ralf Brauksiepe", ] |> head()

```



```{r}
write.csv(bt_speeches_meta, "data/bundestag_cleaned_periods181920.csv")

```
The dataset bundestag_seit2020 contains 22137 speeches from 2020-01-15 to 2023-04-26

The dataset bundestag_181920 contains 57041 speeches from 2013-10-22 to 2023-04-26

Running transformer model PopBERT by luerhard


