---
title: The New Sound of Politics
authors:
  - name: Robert Bowien
    affiliation: University College Dublin
    roles: writing
    corresponding: true
bibliography: references.bib
format: pdf
---

```{r}
library(ggplot2)
library(dplyr)
library(stringr)
```



# Introduction

Watching the prime minister of Bavaria eat sausages and McDonald meals is probably not what citizens from pre-internet Germany imagined political communication would look like in the future. Yet, nowadays politicians like Markus Söder prove that showing approachable personal sides like the consumption of greasy food is a successful way of connecting with the people. Markus Söder has found a way within the set of rules set by the algorithms of social media to connect with the people.

His social media presence features many political messages, but a lot seems non-political and is closer to the appearance of influencer posts. While his food-centered approach is unique in its success, he is by far not the only politician who tries to connect with the people through approachable appearances. Politicians across all parties mix political and seemingly non-political content in their profiles. Ofcourse these posts are just as political as posts with a clear political message.

The currency in democratic politics are votes and the calculation is following: Appealing to the masses by appearing approachable translates to higher reach on social media, higher popularity, higher agenda-setting power and ultimately a higher vote share. This calculation seems to be correct, as Markus Söder has 800 thousand, followers on Instagram, succeeded only by the far-right leader of the AfD Alice Weidel who has over one million followers. Both politicians are often labeled as populists in public discourse and the tactics employed on social media are similar although Weidel does not post lifestyle or food content. The rhetorical tools often resemble each other as they critique the ruling elites (especially during the traffic-light-coalition), establishing an us-vs-them mentality and appealing to the masses in one or another way. The success of right-wing populist politicians does not seem to be random. Moderate and non-populist political actors have struggled to become equally successful utilising social media to communicate and mobilise voters [@mutascu_online_2025; @bobba_likeability_2018].

A recent study of party-related posts on social media in Germany showcases that politicians from parties on the political fringes are more engaged and successful in their social media presence. Posts that are in connection to extremer parties are shown much quicker than posts of parties in the political centre [@verwiebe_digitalisiert_2025]. Political entertainment is just one topic displayed on social media feeds along with comedy, lifestyle, technology, gaming, cooking and more. This way, political information seamlessly reaches broader audiences. This shows the high potential of social media as a political tool and also how important it is to employ tactics that play into the preferences of the algorithms.

After satisfying the need for watching meat-dominated food being consumed, one might slide into the political side of social media. Once that happens, one will inevitably also see another type of posts: Snippets of parliamentary speeches of politicians not only from right-wing parties but from across the whole political spectrum.

Before social media and the internet, parliamentary debates however, have not been this present in peoples everyday life. Who would voluntarily watch debates unfiltered for hours unpaid? Newspapers and TV-news shows have been the filtering lense through which people perceived a curated subset of the words spoken in parliament. Due to social media, most Germans know how the speakers desk of the Bundestag looks like. Snippets of speeches are frequently posted on social media by parties, politicians or other party-affiliated actors [@sternberg_afd-abgeordnete_2026]. Politicians have an incentive to hold speeches from which short snippets that fulfill the conditions the algorithm dictates can be extracted. Although social media has been one of the main drivers of social science research, the evergrowing relevance for it motivates to look at how the rhetoric in political bodies changes once again.

This paper examines how populists talk and how they employ rhetoric tactics to communicate their ideas to the electorate. Examining the rhetoric of populists allows for a deeper understaning of the nature of populism than focusing only on what outcome it produces.

TODO: improve intro to fit research project, less social meda, more populism!!!  

TODO in the end: create a zusammenfassung with logic arrowqs and stuff

## Research Question

The relevance of the possibility of a social-media-fication of parliamentary rhetoric becomes evident when thinking about the purpose of parliamentary debate. If the nature of discourse in parliamentary discussions is changing even more towards a purely performative status of permanent campaigning, it becomes questionable if or how it can still fulfill its function of truthfully informing citizens about politicians actions.

The questions guiding this paper therefore will be following:

*How has parliamentary rhetoric in the German Bundestag changed in recent years? Has it become more populist? And if so, how does that most recent populism sound like?*

How has is social media influencing the tone in parliamentary discussions across Europe? And what does it have to do with populism?

Do politicians adjust their tone in parliamentary speeches to generate more reach on social media? If so, what is the chosen strategy for which parties and why is it populism?

# Literature Review

A common narrative found in European media outlets is that populism and its rhetorical traits have risen during the last years. Some scholars question this narrative and indeed find evidence that the situation is more complicated. In the Netherlands, the occurance of populist rhetoric seems to follow no clear trend. This might be due to self-dooming nature of the anti-establishment rhetoric. After populist parties achieve electoral success, they have joined the establishment and it becomes harder to credibly criticise it [@bos_populist_2014].

This is in line with the claim of populism being endemic in Europe. It has not risen in the last few years, it has long found its way into political rhetoric across all parties [@mazzoleni_socially_2018]. Populist rhetoric is not exclusive to obviously populist figures. It has become a toolbox for a multitude of political figures to choose from. Ofcourse not every politician uses the same rhetoric tools. One goal of this work is to identify how populist rhetoric differs across parties. To do this, it first need to be clear, what populism is.

Defining populism is a near impossible task. Scholars have even referred to this struggle as Cinderella-complex (Aschenputtel-Komplex): There is a shoe formed like populism, but there is no foot fitting the shoe [@priester_rechter_2012]. Nevertheless, one definition has been widely adopted to act as a common baseline understanding of populism; the "ideational" approach. It defines populism as a “thin-centered ideology that considers society to be ultimately separated into two homogeneous and antagonistic groups, ‘the pure people’ versus the ‘corrupt elite,’ and which argues that politics should be an expression of the volonté générale (general will) of the people” [@mudde_populist_2007, p.23]. Populism as thin-centered ideology includes an appeal to "the people" and attacking the elites. An extensive amount of literature relies on these or in their meaning similar concepts [@mazzoleni_socially_2018; @yilmaz_populism_2026; @priester_rechter_2012; @roth_right-wing_2023]. This paper therefore uses these common traits and adds two more in accordance to @erhard_popbert_2025 that enhance the grasp of the concept based on the understanding that populism occurs in the real world often attached to a "host ideology" [@hunger_whats_2022]. This approach has been the base to the definition used by @erhard_popbert_2025 to identify criteria for the training of a NLP transformer model called "PopBert". Additionaly, they add two "thickening" criteria, namely left-wing or right-wing populism. The model therefore rates text on four dimensions: anti-elitism, people-centrism, left-wing host-ideology and right-wing host-ideology.

## Modern populist rhetoric

When populism is attached to another ideology, it "thickens" and becomes what scholars call right-wing or left-wing populism [@hunger_whats_2022]. If the attached ideology is nativism, scholars call it right-wing populism, if it is socialism, scholars call it left-wing populism [@erhard_popbert_2025]. The nature of right-wing populism that is connected to nativism is exclusionary. It's identity creation stems from excludeing non-native concepts and persons. In contrast left-wing populism that is connected to socialism is inclusionary, claiming to represent marginalized or suppressed groups [@erhard_popbert_2025].

Few scholars try to distinguish between types of populist rhetoric. Everything seems to be either populist or not. This lack of conceptual versatility poses a problem since it makes it difficult to understand multi-faceted rhetoric styles. Scholars see populism always presenting a moralistic struggle: The "good people" against the "corrupt" elite [@mudde_populist_2004, @dai_when_2022]. This dynamic fosters a continuing diffusion of what is populist and what is popular behaviour [@mazzoleni_media_2003]. Hence  moralistic struggle is added as a third condition to identify populism in accordance to recent research [@dai_when_2022, @erhard_popbert_2025]. 

## Social media and populism

Before the rise of social media politicians ways of communicating with the electorate were restricted to the rules and filters of media outlets and reporters when communicating with the masses. Media was seen as something the politicians needed to become popular. Nowadays, the relationship shifted. Popular politicians are needed by the media to generate engagement and therefore revenue [@mazzoleni_socially_2018]. Instead of the filtering lenses of journalists, the rules of algorithms now dictates which messages are communicated to the masses.

Since it has become evident that politicians successfully utilise this lacking filter via social media, a lot of research on populist rhetoric focused on and supported this claim [@stieglitz_social_2013; @engesser_populism_2017; @mazzoleni_socially_2018; @mutascu_online_2025; @stavrakakis_research_2024; @bos_populist_2014; @bobba_likeability_2018; @blassnig_populist_2021].

Digital media has introduced a quick and intense change into how we think political information is consumed by the electorate. Consuming political online content can contribute to political participation [@kahne_political_2018; @boulianne_social_2015]. Political communication on social media seems to benefit right-wing parties more than left-wing parties [@mutascu_online_2025]. Political communication that employs populist rhetoric on social media differs across parties. Left-wing populists prefer attacks on the economic elite while right-wing populists predominantly attack the media elite and ostracise outgroups [@engesser_populism_2017].



Social media performance plays an increasingly important role in parties communication strategies. Parties pay attention to the performance of their social media posts and the engagement with them and adjust their strategy accordingly [@ennser-jedenastik_does_2022]. Since social media proved to be an effective tool for influencing public opinion, political advisors have put more salience on social media [@lilleker_towards_2015]. It is therefore reasonable to assume that parties try to increase their social media reach, for example by using populist rhetoric in speech snippets from parliamentary bodies.


```{r}
# data_sm_relevance <- read.csv("data/sm_relevance.csv")
data_sm_relevance <- readRDS("C:/Users/rober/UCD/pol42050/new-sound-of-politics/data/sm_relevance.rds")
```
```{r}

# Compute share of each source within age group x wave
df_counts <- data_sm_relevance |>
  group_by(time_period, age_group, media_info_source) |>
  summarise(n = n(), .groups = "drop") |>
  group_by(time_period, age_group) |>
  mutate(share = n / sum(n)) |>
  ungroup()

cbPalette <- c("#E69F00", "#009E73",  "#D55E00", "#56B4E9", "#F0E442", "#0072B2", "#CC79A7")

# Colour palette per source
source_palette <- data.frame(media = c("TV",
  "Newspaper",
  "Radio",
  "Social media",
  "Other Internetproviders",
  "Personal conversation",
  "Other source"),
  colour = cbPalette
)

# create working colour palette
source_palette <- setNames(source_palette$colour, source_palette$media)
 
# Plot — faceted by age group, one line per source
p1 <- ggplot(df_counts, aes(x = time_period, y = share, colour = media_info_source, group = media_info_source)) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2) +
  facet_wrap(~ age_group, ncol = 1) +
  scale_x_date(
    date_labels = "%b %Y",
    date_breaks = "6 months",
    expand = expansion(mult = 0.03)
  ) +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, NA)
  ) +
  scale_colour_manual(values = source_palette, name = "Information source") +
  labs(
    title    = "Information source by age group over time",
    subtitle = "Share of respondents per source, split by age groups",
    x        = "Survey period (start)",
    y        = "Share"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title       = element_text(face = "bold", size = 14),
    plot.subtitle    = element_text(colour = "grey40", margin = margin(b = 8)),
    legend.position  = "right",
    axis.text.x      = element_text(angle = 30, hjust = 1),
    panel.grid.minor = element_blank(),
    strip.text       = element_text(face = "bold", size = 12)
  )
 
# Save
# ggsave("plot_sources_by_agegroup.png", plot = p1, width = 11, height = 9, dpi = 150)
p1

```
As illustrated in figure 1, social media has established itself as the second most important source of political information. For young people between 18 and 29 it even is the most important source. 

TDOD: Work this into paper: IF FINDINGS SHOW AfD and LINKE rhetoric adjusts to be more populist for social media, this might be one explanation for their recent success within young voting groups.

Furthermore, existing research demonstrates that populist rhetoric is widely utilised by political actors through social media [@mutascu_online_2025; @engesser_populism_2017; @mazzoleni_socially_2018]. In a German context, a recent study on the utilisation of social media by political actors revealed that emotionally laden topics perform better than neutral policy discussions [@mersmann_digitization_2026].

## The research gap

While most studies focus on how social media is used to communicate to the public, fewer scholars aimed at the core of democratic discourse: the parliament. The parliament is a central institution of every democracy. Its stated purpose is to provide meaningful discussion open to the public to keep the electorate informed and able to hold politicians accountable. Previously, this discourse was filtered through the lense of media actors. With the social media enabling politicians to circumvent this filter, shedding light on potential changes in discourse culture increases in relevance.

Foregoing review of the literature mentioned a link between the relevance of social media as political information medium and the rhetoric patterns in parliamentary speeches. It has opened a door to examine and understand rhetoric changes in parliament. This will be elaborated on in the following section on theoretical considerations.

# Theory

For political actors it is of crucial importance to be able to effectively communicate their ideas and messages to the electorate. In a quickly changing environment where social media is omnipresent and supports direct, raw and agitating messages, political actors are incentivised to develop new ways of communication that adhere to the rules of the algorithm in some way. The reviewed literature suggests that populism offers successful tools to achieve this.

It is important to understand that the theoretical link here relies not on the assumption that the electorate pays attention to speeches in the Bundestag directly. Referring back to the literature, the electorate rarely ever listens to speeches, but rather relies on other media as filtering instance to inform themselves [@blumler_third_1999]. This work assumes political actors prioritise social media to speak to the electorate [@lilleker_towards_2015]. Instead of or additionaly to creating content specifically for social media, they adjust Bundestag speeches to fit social media rules. Topics that do well on social media are those that fit well into the populist playbook. Emotionally laden and 

It is easier to critizise the ruling elites when you are in the opposition. In Germany, the parties known for populist rhetoric are the far left and far right parties which are not in power. We can expect them to use populist rhetoric to a greater extent as they do not have governing power. To therefore conclude that parties of the political centre do not use populist rhetoric would be wrong. Examples of elements of populist rhetoric can be examined across all parties. Research on populism on Italian facebook finds that populist rhetoric is endemic across party boundaries [@mazzoleni_socially_2018]. Political leaders try to appeal to the masses, some more than others, often by pop-cultural ways. From "down to earth" asthetics campaigning of Robert Habeck in the latest parliamentary election to professionally unprofessional food blogging by Markus Söder, politicians find a way to make the masses feel close to that charismatic political figure.

Hence, it seems that populist rhetoric is employed both in and out of government and even though it takes on different forms, no politician seems to be able to escape it. An advantage of populist rhetoric lies in its ability to always offer the solution that nobody else offers. To be most credible with this unique solution, it is best to present yourself further away from the mainstream, where you can easily come up with something that is not commonly talked about, something that is unique and plays into the rules of algorithms. This should be easiest for opposition and extreme parties, which is why it can be expected that they were the first to adjust their rhetoric. With the ongoing success, moderate parties should eventually try to utilise populist tools too.

Building on this, if speakers care about their potential social media performance when delivering a speech, they should engage more in populist rhetoric when addressing emotionally laden topics.  


## Hypotheses

Based on preceding literature review and theoretical considerations, following hypotheses are specified:

Hypothesis 1: The scores for populist rhetoric in parliamentary speeches will increase over time.

As social media becomes more relevant as a medium of political information, we would expect members of parliament to adjust their rhetoric to fit social media criteria.

H2a: Parties in opposition are expected to use more anti-elitism rhetoric than governing parties.

This is straightforward: Being in government makes it more difficult to critizise elites as this means critizising yourself. 

H2b: Extreme parties in opposition should show the highest level of populist rhetoric with left-wing parties showing a higher emphasis on people-centrism and right-wing parties on anti-elitism.

H3: The scores for populist rhetoric increases when speakers adress topics that are prone to carry populist rhetoric.



# Methodology

It is still contested and context-dependent how populist rhetoric should be operationalised [@popping_measuring_2018]. Before the availability of transformer models, traditional natural language processing (NLP) often employed dictionary based methods like bag-of-words approaches. These approaches are very reliable and reproducable but face significant challenges when it comes to complex lingual patterns. Populist rhetoric is full of various linguistic tricks which often are difficult to define and capture with dictionary based approaches. Furthermore, dictionaries designed to capture populism often differ and it is not clear when a dictionary is extensive [@pauwels_chapter_2017]. Transformer models offer a promising approach to successfully identify populist rhetoric by relying on word embeddings, which capture the context in which a word is used [@erhard_popbert_2025, @bonikowski_politics_2022].

Recent research on German populist rhetoric has employed a BERT (Bidirectional Encoder Representations from Transformers) model and fine-tuned it to avoid dichotomous classifying of populism and rather classify aspects of their communications [@erhard_popbert_2025]. 
This paper follows a similar approach in many aspects, but it differs in small, but relevant details. @erhard_popbert_2025 used PopBERT, a fine tuned version of the GBERT model, which is the German version of BERT. Since the model has already been trained on parliamentary speeches from the Bundestag and is fine tuned to identify populist rhetoric, no fine-tuning is needed anymore. It can be directly applied to the more recent data this paper aims to examine. In accordance with @erhard_popbert_2025 the model will classify populist rhetoric on sentence level

The dataset is constructed of speech snippets in the German Bundestag. A speech is defined by the creators of the dataset as "...all verbal acts delivered by an actor during a single plenary session." One speech can be reconstructed by grouping the data by speaker, session, date and (... TODO: exact description of coding!). However, the data are not perfectly clean. Some snippets include meta data and random spots often mid-word, e.g.: "[...]Der Abgeordnete Dr. Norbert Lammert ist vorge Deutscher Bundestag – 18. Wahlperiode – 1. Sitzung. Berlin, Dienstag, den 22. Oktober 2013 5 Alterspräsident Dr. Heinz Riesenhuber" where the next snippet begins with " schlagen worden. – Ich sehe keine weiteren Vorschläge, ich höre keinen Widerspruch.[...]". To clean this as good as possible, a removing regex pattern has been applied to the corpus before splitting the speeches into sentences. Furthermore, all snippets that are interjections and not part of a speak and snippets where the speaker could not be identified are removed from the dataset. This leaves us with 44501 sentences or 6674 speeches from October 2021 to April 2023.

Furthermore, it occurs that speeches are wrongly attributed to the chair rather than the speaker. Since it is not possible to determine how often this happens, any analysis conducted by speaker or party would be meaningless. To still be able to analyse the rhetoric by party and speaker, all speeches that are attributed to the chair are excluded. While this reduces the sample size and therefore statistical power, the quality of the data is maintained.

This paper does not classify the examined speakers as populist or non-populist actors, it classifies single sentences as populist on 4 possible dimensions. It relies on PopBERT to identify populist rhetoric across all types of speakers since the goal of this work is to identify differences in language over time and between parties and not to determine dichotomously who is populist and who is not.

The SpeakGer speech corpus is used because it is the most recent and easily accessible dataset on Bundestag speeches. It includes all speeches from 1947 until 2023. This work focuses on the most recent complete period for the 20th Bundestag from October 2021 until April 2023, which was not examined by @erhard_popbert_2025. Furthermore, speeches from speakers who were not identified and all interjections are excluded.
Since the original model was trained and validated on Bundestag speeches originally, no fine-tuning or additional validation was conducted because the language should not change strongly over the additional period that is now focus of the data.

Although extensive measures to clean the corpus have been undertaken, some limitations remain. Whenever there is a parliamentary interview, both the question and the answer are treated as speeches by the questioner. In some cases this could influence the amount of populism attributed to one speaker. However, this bias should be random since both questioner and interviewed could be employing populist rhetoric. The effort to remove those parts would probably exceed the benefit they pose for data quality.

Another caveat is the data quality provided by the dataset. @lange_speakger_2023 sometimes provide incomplete speeches. E.g. the speech of MP Florian Hahn on the 2013-12-18 Only includes the first few sentences^[Mr. Hahn's speech was more than 800 words long. The speech in the dataset: "Sehr geehrte Frau Präsidentin! Kolleginnen und Kol legen! Die Grosse Koalition hat sich gefunden. Wir haben gestern die Kanzlerin gewählt. Es hat sich eine neue, gute Regierung gebildet. „Endlich“, werden die meisten sagen, man könnte aber auch sagen: Gut Ding braucht Weile. – Mit Blick nach Hessen, wo die Regierung noch" [@noauthor_plenarprotokoll_2013]], although the whole speech actually was much longer. Best practice to have the most accurate and complete dataset would be to not rely on the dataset of @lange_speakger_2023, but directly scrape the documents from "https://www.bundestag.de/dokumente/protokolle/plenarprotokolle". However, these efforts would be exceeding the authors capabilites.

To understand the differences on how political topics are discussed in the Bundestag and on social media topic modelling is used. The topics are derived from unsupervised modelling and where it is appropriate, the same topics as in @mersmann_digitization_2026 are used to classify text.


Check in how many speeches the word gendern occurs across parties in the Bundestag. --\> Hypothesis: populist parties like the AfD and CSU talk a lot about them, others do not. \<-- They set the agenda and aim to make it look like big issue for the other (left-green) parties. Could also do a relative frequency analysis using slides from week 5 (chi²/Keyness analysis). expectation: populist parties will talk more about populist topics lol

```{r}
# summary table key variables
keyvars <- data.frame("Variable" = c("Populist dimension", "Populist dimension", "Populist dimension", "Populist dimension", "Topic", "Party Affiliation", "Speaker", "Time"),
"Type" = c("Ratio", "Ratio", "Ratio", "Ratio", "Ratio", "Nominal", "Nominal", "Interval"),
"Operationalisation" = c("Anti-Elitism", "People-Centrism", "Left-Wing Host-Ideology", "Right-Wing Host-Ideology", "Topic proportions", "Speaker party", "Speaker name", "monthly intervals (Oct 2021 - Apr 2023)")) #TODO explain why time is continuous (change over time is expected to be linear and not bound by events)

print(keyvars)

```

# Results and Discussion






Goal: Show how populist rhetoric developed over time (from 2015)



TODO: interesting to check Die Linke since they split into Linke and BSW.

## Limitations

The level of populism is evaluated by PopBERT on sentence instead of speech level. This poses two possible problems:

First, the available data from SpeakGer comes on speech level. Splitting the speeches into sentences may wrongly split sentences at spots where they should not be split and vice versa. Parliamentary often are no cleanly written texts. During a speech there might be interventions or slips of the tongue that lead to difficulties in determining the end and beginning of sentences.
Second, classifying on speech level would prevent missing any context information. Speech-level analysis focuses on the sentiment expressed throughout the whole text, whereas sentence-level analysis is normally employed when concentrating on short messages [@aliyu_comparative_2024]. Two sentences on their own might not be enough to identify populist rhetoric, but when they are taken together, they might prove to be populist.

However, the PopBERT is limited to 512 tokens per classification, which would require chunking speeches to avoid truncating them. This would keep more context information but the chunk might split mid-sentence which would cause again loss of information. Therefore, this work relied on classifying at sentence level, which is closest to the method used by @erhard_popbert_2025. For future research, it might be valuable to employ more resource intesive approach to enable classifying on speech instead of sentence level. 

Robustness checks for the transformer model have not been sufficiently conducted. It relies on the robustnesschecks of @erhard_popbert_2025 which have been conducted on data of the same institution in the same context but is older data. It is therefore not ensured that it performs the same on data that is more recent. This is problematic because this work explicitly aims to examine how populist rhetoric changes over time. Using a model trained on speeches from an earlier time period might reduce its performance (TODO: what metrics may be reduced?). However, using the same model ensures that any differences identified do not come from any differences in the transformer model. 

To examine a direct link between the relevance of social media and the amount of populist rhetoric in the German parliament, a multi-level regression of survey data on the populist score controlling for factors such as party, speaker and time would be necessary. That would rely on an instable framework of assumptions and is therefore left out from this paper.
This paper is limited to data on parliamentary speeches and unrelated social media metrics. To consolidate inferences about a change in parliamentary rhetoric due to the increased relevance of social media, more data is needed. Matching social media posts with speeches in a quantitative rather than limited descriptive manner (like in this paper) would provide a clear image of which and when specific rhetoric is employed in parliament. 

Furthermore, this work does not address the demand side of populist/popular rhetoric on social media. To understand, first, in which cases populist/popular rhetoric generates reach successfully, and second, in which cases digital success translates into political success, the demand side is a crucial, yet under-researched topic.

The survey data has been collected online. This might bias the data towards an audience that is prone to use digital information and therefore overestimate the overall relevance of social media for the electorate.

# Conclusion

For future research, it should be considered to use LLMs to identify populist rhetoric because they perform better when it comes to identifying nuanced stances. 
Future work on the utilisation of social media by political actors should take into consideration not only accounts clearly representing a specific party. Accounts that are not official accounts but still push narratives close to a certain parties perspective might have considerably more influence than the political accounts themselves. Here, whole networks should be the subject of research.

# NOTES

Look for words that may come from social media culture

look for complexity of speech, social media wants easy stuff, not complicated long sentences, attention span is getting shorter, etc.

