---
title: The New Sound of Politics
author:
  - name: Robert Bowien
  - affiliation: University College Dublin
  - module: Quantitative text analysis
  - coordinator: Prof. Dr. Stefan Müller
    roles: writing
    corresponding: true
bibliography: references.bib
fig-pos: "H"
tbl-pos: "H"
execute:
  echo: false
output:
  pdf_document:
    extra_dependencies: ["longtable", "float"]
    toc: true
---

```{r}
#| label: setup
library(ggplot2)
library(dplyr)
library(stringr)
library(knitr)
library(kableExtra)
library(quanteda)
library(quanteda.textplots)
library(patchwork)
library(huxtable)
library(lme4)
library(keyATM)

data_sm_relevance <- readRDS("C:/Users/rober/UCD/pol42050/new-sound-of-politics/data/sm_relevance.rds")
bt_speeches_long <- readRDS("C:/Users/rober/UCD/pol42050/new-sound-of-politics/data/bt_speeches_long.rds")
keyness_by_party <- readRDS("C:/Users/rober/UCD/pol42050/new-sound-of-politics/data/keyness_by_party3.rds")
keyness_by_party <- lapply(keyness_by_party, function(x) {
  class(x) <- c("keyness", "textstat", "data.frame")
  x
})
m_ae <- readRDS("C:/Users/rober/UCD/pol42050/new-sound-of-politics/data/m_ae.rds")
m_pc <- readRDS("C:/Users/rober/UCD/pol42050/new-sound-of-politics/data/m_pc.rds")
m_lw <- readRDS("C:/Users/rober/UCD/pol42050/new-sound-of-politics/data/m_lw.rds")
m_rw <- readRDS("C:/Users/rober/UCD/pol42050/new-sound-of-politics/data/m_rw.rds")
keyatm_model    <- readRDS("C:/Users/rober/UCD/pol42050/new-sound-of-politics/data/keyatm_model_digitization3.rds")
fig_speaker_pop <- readRDS("C:/Users/rober/UCD/pol42050/new-sound-of-politics/data/fig_speaker_pop.rds")
fig_topic_party <- readRDS("C:/Users/rober/UCD/pol42050/new-sound-of-politics/data/fig_topic_party_pop.rds")
fig_time_trends <- readRDS("C:/Users/rober/UCD/pol42050/new-sound-of-politics/data/fig_time_trends.rds")
topic_coef_labels <- readRDS("C:/Users/rober/UCD/pol42050/new-sound-of-politics/data/topic_coef_labels.rds")
topic_label_map   <- setNames(topic_coef_labels$nice_name, topic_coef_labels$coef_name)

bundestag_palette <- c(
  "CDU/CSU" = "#000000", "CDU" = "#333333", "CSU" = "#555555",
  "SPD"     = "#E3000F", "AfD" = "#009EE0", "FDP" = "#FFED00",
  "B90/Die Grünen" = "#1FA12D", "Die Linke" = "#BE3075", "BSW" = "#9B4A97"
)

theme_set(theme_minimal())
options(knitr.table.format = "latex")
options(huxtable.latex_float = "H")

```

# Abstract

This work explores populist parliamentary rhetoric in the context of social media as a potential influental factor. It is theorised that politicians aim to increase their reach on social media as the electorate increasingly obtain political information from this medium. This relation is explored descriptevely and inferentially. PopBERT is employed to classify populist rhetoric and topic modelling identifies prevalent topics in speeches. A Logistic generalized linear mixed effects regression model is used to estimate effects of time, topic, and party affiliation on the amount of populist rhetoric. While no inferential conclusions can be drawn for the influence on social media on parliamentary rhetoric, party affiliation and specific topics are found to have significant effects.


# Introduction

Before social media and the internet, parliamentary debates have not been present in peoples everyday life. Who would voluntarily watch debates unfiltered for hours unpaid? Newspapers and TV-news shows have been the filtering lense through which people perceived a curated subset of the words spoken in parliament. Due to social media, many Germans nowadays know exactly how the speakers desk of the Bundestag looks like. Snippets of speeches are frequently posted on social media by parties, politicians or other party-affiliated actors [@sternberg_afd-abgeordnete_2026].

Modern day populism is closely accompanied by social media. Social media bypasses traditional forms of communication between politicians and the electorate. [@blumer_third_1999; @engesser_populism_2017; @stavrakakis_research_2024; @mutascu_online_2025]. The faces of the most popular politicians can be found all over social media, communicating more directly to voters than ever before.

A recent study of party-related posts on social media in Germany showcases that politicians from parties on the political fringes are more engaged and successful in their social media presence. Posts that are in connection to extremer parties are shown much quicker than posts of parties in the political centre [@verwiebe_digitalisiert_2025]. Political entertainment is just one topic displayed on social media feeds along with comedy, lifestyle, technology, gaming, cooking and more. This way, political information seamlessly reaches broader audiences. One type of posts is especially interesting: Snippets of parliamentary speeches.

## Research Question

The relevance of the possibility of a social-media-fication of parliamentary rhetoric becomes evident when thinking about the purpose of parliamentary debate. If the nature of discourse in parliamentary discussions is changing towards a performative status of permanent campaigning, it becomes questionable to what extent it fulfills its function of truthfully informing citizens about politicians actions.

The questions guiding this paper will be following:

*Could the way populist rhetoric is employed be influenced by politicians goal to create reach on social media?*

*How does populist rhetoric differ across parties?*

*How is the topic of a speech influencing the amount of populist rhetoric in Bundestag speeches?*

# Literature Review

A common narrative found in European media outlets is that populism and its rhetorical traits have risen during the last years. Some scholars question this narrative and indeed find evidence that the situation is more complicated. In the Netherlands, the occurance of populist rhetoric seems to follow no clear trend. This might be due to self-dooming nature of the anti-establishment rhetoric. After populist parties achieve electoral success, they have joined the establishment and it becomes harder to credibly criticise it [@bos_populist_2014].

This is in line with the claim of populism being endemic in Europe. It has not risen in the last few years, it has long found its way into political rhetoric across many parties [@mazzoleni_socially_2018]. Populist rhetoric is not exclusive to obviously populist figures. It has become a toolbox for a multitude of political figures to choose from.

Defining populism is a near impossible task. Scholars have even referred to this struggle as Cinderella-complex (Aschenputtel-Komplex): There is a shoe formed like populism, but there is no foot fitting the shoe [@priester_rechter_2012]. Nevertheless, one definition has been widely adopted to act as a common baseline understanding of populism; the "ideational" approach. It defines populism as a “thin-centered ideology that considers society to be ultimately separated into two homogeneous and antagonistic groups, ‘the pure people’ versus the ‘corrupt elite,’ and which argues that politics should be an expression of the volonté générale (general will) of the people” [@mudde_populist_2007, p.23].

Populism as thin-centered ideology includes an appeal to "the people" and attacking the elites. An extensive amount of literature relies on these or in their meaning similar concepts [@mazzoleni_socially_2018; @yilmaz_populism_2026; @priester_rechter_2012; @roth_right-wing_2023]. This paper therefore uses these common traits and adds to it based on the understanding that populism occurs in the real world often attached to a "host-ideology" [@hunger_whats_2022].

## Social media and populism

Consuming political content online can not only shift opinions, but also contribute to political participation [@kahne_political_2018; @boulianne_social_2015]. Hence, parties pay attention to the performance of their social media posts and adjust their strategy accordingly [@ennser-jedenastik_does_2022; @lilleker_towards_2015].

Politicians ways of communicating with the electorate used to be restricted to the rules and filters of media outlets and reporters when communicating with the masses. Nowadays, the relationship shifted. Instead of the filtering lenses of journalists, the rules of algorithms now dictates which messages are communicated to the masses.

Since it has become evident that politicians successfully utilise this lacking filter via social media, a lot of research on populist rhetoric focused on and supported this claim [@stieglitz_social_2013; @engesser_populism_2017; @mazzoleni_socially_2018; @mutascu_online_2025; @stavrakakis_research_2024; @bos_populist_2014; @bobba_likeability_2018; @blassnig_populist_2021].

Examples of elements of populist rhetoric can be examined across all parties. Research on populism in Italian facebook posts finds that populist rhetoric is endemic across party boundaries [@mazzoleni_socially_2018]. In a German context, a recent study on the utilisation of social media by political actors revealed that emotionally laden topics perform better than neutral policy discussions [@mersmann_digitization_2026].

## Relevance

Most studies on effects of social media on the political public. Few ask how social media influences the political domain. This work aims at the parliament. After creating legislation, an essential function is to provide meaningful discussion open to the public to keep the electorate informed and able to hold politicians accountable. Potential systematic external influence on the discourse due to the need to perform well on social media can have hindering effects on the parliamentary capabilities.

# Theory

The theoretical link relies not on the assumption that the electorate pays attention to speeches in the Bundestag directly. This work instead assumes political actors prioritise social media to speak to the electorate [@lilleker_towards_2015]. To support their digital appearance, speakers might adjust their rhetoric to adhere to social media algorithms. Topics that do well on social media are those that fit well into the populist playbook. Addressing emotionally laden and controversial topics should be advantageous for the employment of populist rhetoric.

## Hypotheses

Based on preceding literature review and theoretical considerations, following hypotheses are specified:

*Hypothesis 1*: The scores for populist rhetoric in parliamentary speeches will increase over time.

As social media becomes more relevant as a medium of political information, we would expect members of parliament to adjust their rhetoric to fit social media criteria.

*Hypothesis 2*: Extreme parties should show the highest level of populist rhetoric with left-wing parties showing a higher emphasis on people-centrism and right-wing parties on anti-elitism.

The success of right-wing populist politicians does not seem to be random. Moderate and non-populist political actors have struggled to become equally successful utilising social media to communicate and mobilise voters [@mutascu_online_2025; @bobba_likeability_2018]. This difference might be due to different tactics: Anti-elitism is expected to be employed more intensively by right-wing parties, whereas left-wing parties should show a higher degree of people-centrism [@mudde_populist_2004].

*Hypothesis 3*: Populist rhetoric should be higher for topics that are emotionally laden.

If speakers care about their potential social media performance when delivering a speech, they should engage more in populist rhetoric when addressing emotionally laden topics.

# Methodology

It is still contested and context-dependent how populist rhetoric should be operationalised [@popping_measuring_2018]. Before the availability of transformer models, traditional natural language processing (NLP) often employed dictionary based methods like bag-of-words approaches. These approaches are easily reproducable but face significant challenges when it comes to complex lingual patterns. Populist rhetoric is full of various linguistic tricks which often are difficult to define and capture with dictionary based approaches. Furthermore, dictionaries designed to capture populism often differ and it is not clear when a dictionary is extensive [@pauwels_chapter_2017].

This work instead relies on a transformer model called PopBERT. Transformer models offer a promising approach to successfully identify populist rhetoric by relying on word embeddings, which capture the context in which a word is used [@erhard_popbert_2025, @bonikowski_politics_2022].  It classifies text on four dimensions: anti-elitism, people-centrism, left-wing host-ideology and right-wing host-ideology.

The model was trained and validated on Bundestag speeches from the 18th and 19th period by @erhard_popbert_2025. They employed a BERT (Bidirectional Encoder Representations from Transformers) model and fine-tuned it.

To understand the differences on how political topics are discussed in the Bundestag topic modelling is used. The topics are derived from iterated unsupervised modelling and where it is appropriate, the same topics as in @mersmann_digitization_2026 have been adopted.

The corpus was collected using the official API of the "Dokumentations- und Informationssystems für Parlamentsmaterialien" (DIP) [@noauthor_deutscher_nodate]. After preprocessing, it consists of 23163 speeches and 1168085 sentences from the most recent complete Bundestag period between October 2021 and April 2023.

Four mixed effects logistic regression models are calculated to examine the effect of the key variables (table 2) on the PopBERT predictions. This is suitable for modelling binary dependent variables where independent variables have fixed and random effects [@noauthor_mixed_nodate].

Sentences of a speech that has been classified as mostly procedural were excluded from the analysis to prevent them from deflating populist rhetoric counts.

```{r}
#| label: tbl-results
#| tbl-cap: "regression key variable table"

keyvars <- data.frame("Variable" = c("Anti-Elitism", "People-Centrism", "Left-Wing Host-Ideology", "Right-Wing Host-Ideology", "Topic", "Time", "Party Affiliation", "Speaker"),
"Type" = c("Binary", "Binary", "Binary", "Binary", "Ratio", "Interval", "Nominal", "Nominal"),
"Operationalisation" = c("PopBERT prediction", "PopBERT prediction", "PopBERT prediction", "PopBERT prediction", "Topic proportions", "Monthly intervals", "Speaker party", "Speaker name"),
"Theoretical function" = c("Dependent variable", "Dependent variable", "Dependent variable", "Dependent variable", "Independent variable", "Independent variable", "Independent variable", "Control variable"))


keyvars |>
  kable(
    booktabs = TRUE,
    longtable = FALSE,
    align = c("l", "l", "l", "l"),
    escape = FALSE
  ) |>
  kable_styling(
    latex_options = c("HOLD_position", "scale_down"),
    font_size = 9,
    full_width = TRUE
  ) |>
  column_spec(1, width = "6.5cm") %>%
  column_spec(2, width = "2cm") %>%
  column_spec(3, width = "5cm") %>%
  column_spec(4, width = "4cm") %>%
  row_spec(0, bold = TRUE)  # Bold headers


```

# Results and Discussion

Of all 1177428 sentences classified by PopBERT, 9.4% contian anti-elitism, 2.3% contain people-centrism, 1.2% have a left-wing host ideology and 0.9% have a right-wing host ideology.^[These proportions are close to the results of @erhard_popbert_2025, p.8: "...we find 7.4% to contain antielitism, 2.1% people-centrism, 1.2% a left-wing host ideology, and 0.5% a right-wing host ideology."]

## Hypothesis 1

Figure 1 highlights the relevance of social media as political information medium. It has established itself as the second most important source of political information during the times of the traffic light coalition. For young people between 18 and 29 it even is the most important source.

```{r}
#| label: fig-sm_over_time
#| fig-cap: Political information sources over time
#| fig-width: 12
#| fig-height: 8
# Compute share of each source within age group x wave
df_counts <- data_sm_relevance |>
  group_by(time_period, age_group, media_info_source) |>
  summarise(n = n(), .groups = "drop") |>
  group_by(time_period, age_group) |>
  mutate(share = n / sum(n)) |>
  ungroup()

cbPalette <- c("#E69F00", "#009E73",  "#D55E00", "#56B4E9", "#F0E442", "#0072B2", "#CC79A7")

# Colour palette per source
source_palette <- data.frame(media = c("TV",
  "Newspaper",
  "Radio",
  "Social media",
  "Other Internetproviders",
  "Personal conversation",
  "Other source"),
  colour = cbPalette
)

# create working colour palette
source_palette <- setNames(source_palette$colour, source_palette$media)
 
# Plot — faceted by age group, one line per source
p1 <- ggplot(df_counts, aes(x = time_period, y = share, colour = media_info_source, group = media_info_source)) +
  geom_line() +
  geom_point(size = 2) +
  facet_wrap(~ age_group, ncol = 1) +
  scale_x_date(
    date_labels = "%b %Y",
    date_breaks = "6 months"
    # expand = expansion(mult = 0.03)
  ) +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, NA)
  ) +
  scale_colour_manual(values = source_palette, name = "Information source") +
  labs(
    title    = "Information source by age group over time",
    subtitle = "Share of respondents per source, split by age groups",
    x        = "Survey period (start)",
    y        = "Share"
  )
p1

```

Figure 2 explores the development of the four dimensions of populist rhetoric and social media as a political information medium.

```{r}
#| label: fig-sm-vs-populism
#| fig-cap: "Mean populism rate in Bundestag speeches by dimension (solid coloured lines, left axis) and share of GLES respondents citing social media as primary political information source (grey dashed line, right axis), aggregated by quarter and survey wave respectively."
#| fig-height: 6
#| fig-width: 12
sm_scale  <- fig_time_trends$sm_scale
sm_offset <- fig_time_trends$sm_offset

dim_palette <- c(
  "Anti-Elitism"              = "#E69F00",
  "People-Centrism"           = "#009E73",
  "Left-Wing Host-Ideology"   = "#56B4E9",
  "Right-Wing Host-Ideology"  = "#F0E442",
  "Social media share"        = "grey50"
)

ggplot() +
  geom_line(
    data = fig_time_trends$pop,
    aes(x = quarter, y = mean_score, colour = dim_label),
    linewidth = 0.8
  ) +
  geom_line(
    data = fig_time_trends$sm,
    aes(x = time_period, y = sm_scaled, colour = "Social media share"),
    linetype = "dashed", linewidth = 0.8
  ) +
  geom_point(
    data = fig_time_trends$sm,
    aes(x = time_period, y = sm_scaled, colour = "Social media share"),
    size = 2
  ) +
  scale_colour_manual(
    values = dim_palette,
    breaks = names(dim_palette),
    name   = NULL
  ) +
  scale_y_continuous(
    name = "Mean populism rate",
    sec.axis = sec_axis(
      ~ (. - sm_offset) / sm_scale,
      name   = "Social media share",
      labels = scales::percent_format(accuracy = 1)
    )
  ) +
  scale_x_date(date_labels = "%b %Y", date_breaks = "6 months") +
  labs(x = NULL) +
  theme(
    axis.text.x     = element_text(angle = 30, hjust = 1),
    legend.position = "bottom"
  )
```


While every dimension of populism increases over time, the increase is most noticable for anti-elitism. This hints at how strong critizising rhetoric is a preferred tool of the populist repertoire. While this supports H1, drawing conclusions requires an inferential analysis of populist rhetoric in speech snippets on social media, especially in regards to anti-elitism.
Unfortunately, the available survey data does not fully cover the 20th period of the Bundestag. This limits the invormativeness of the graph further.

## Hypothesis 2:

```{r}
#| label: fig-speaker-pop
#| fig-cap: "Top 2 most and least populist speakers per party by mean PopBERT score across all four dimensions (minimum 100 classified sentences)."
#| fig-height: 10
#| fig-width: 12
#| 
ggplot(fig_speaker_pop, aes(x = mean_pop, y = Name, colour = Party)) +
  geom_vline(
    aes(xintercept = median_pop),
    linetype = "dashed", colour = "grey50", linewidth = 0.5
  ) +
  geom_segment(
    aes(x = 0, xend = mean_pop, y = Name, yend = Name),
    colour = "grey78"
  ) +
  geom_point(size = 3) +
  scale_colour_manual(values = bundestag_palette, guide = "none") +
  facet_wrap(~ Party, scales = "free_y", ncol = 2) +
  labs(x = "Mean populism score", y = NULL,
       caption = "Dashed line: party median (speakers with ≥100 classified sentences).") +
  theme(
    strip.text   = element_text(face = "bold"),
    axis.text.y  = element_text(size = 8),
    plot.caption = element_text(size = 7, colour = "grey40")
  )
```

Figure 3 yields expected results. Parties on the fringes of the political spectrum show higher amounts of populist rhetoric in accordance to the specified hypotheses.

## Hypothesis 3:

@mersmann_digitization_2026 have applied topic modelling to a corpus of social media posts of political parties and politicians and measured the performance of those posts. When modelling similar topics on the Bundestag speech corpus and measuring the level of populist rhetoric per topic, some parallels emerge. Social media postings that address emotional topics such as immigration policy, security issues, and climate change perform best (@mersmann_digitization_2026). These are also the topics that experience high levels of populist rhetoric in parliamentary debates, both on the anti-elitism and the people-centrism scales. Right wing populism is more/less prevalent than left wing populism.

```{r}
#| label: fig-topic-party-heatmap
#| fig-cap: "Mean populism rate per topic and party, faceted by PopBERT dimension. Higher values indicate a greater share of sentences in that topic–party combination classified as populist."
#| fig-height: 12
#| fig-width: 14
top5_labels <- fig_topic_party |>
  group_by(dim_label) |>
  slice_max(mean_score, n = 10) |>
  ungroup() |>
  mutate(label = scales::percent(mean_score, accuracy = 0.1))

ggplot(fig_topic_party, aes(x = Party, y = topic_label, fill = mean_score)) +
  geom_tile(colour = "white", linewidth = 0.4) +
  geom_text(
    data     = top5_labels,
    aes(label = label),
    colour   = "black",
    size     = 2.5,
    fontface = "bold"
  ) +
  scale_fill_gradient(
    low    = "white",
    high   = "#C0392B",
    name   = "Mean\npopulism\nrate",
    labels = scales::percent_format(accuracy = 1)
  ) +
  facet_wrap(~ dim_label, ncol = 2) +
  labs(x = NULL, y = NULL) +
  theme(
    axis.text.x  = element_text(angle = 45, hjust = 1, size = 7),
    axis.text.y  = element_text(size = 7),
    strip.text   = element_text(face = "bold"),
    panel.grid   = element_blank()
  )
```

Figure 4 supports H3. The main driver for populist rhetoric seems to be the party affiliation. However, some topics seem to funnel more populist rhetoric than others. For Anti-Elitism, party affairs show high levels of populism, especially among parties at the political fringes (AfD, Die Linke, BSW). For people-centrism, migration exhibits the highest levels of populist language across party boundaries. Right-wing sentences employ populist rhetoric more often for topics that are emotionally charged. Migration and Security exhibit the highest mean populism rates for that cluster. Left-wing sentences seem to be dominated by party affiliation. Die Linke holds most of the top 10 percentages here.


A keyness analysis (see appendix) reveals no surprising results. The far-right AfD uses nationalistic language and terms related to the people such as "bürger" or "deutschland" more often while the far-left Die Linke uses terms related to low-income and social policies such aus "mietwucher" and "resilienzbonus" more often than other parties.

## Regression analysis

```{r}
#| label: reg-table

# summary()$coefficients gives fixed-effects only with model-based SEs.
# Wrapping as coeftest class lets huxreg dispatch via broom::tidy.coeftest.
# lmtest::coeftest() itself fails on glmerMod because coef.merMod returns
# all conditional modes (~700 speakers), making diag() hit R's vector length limit.
fe_coeftest <- function(m) {
  cf <- summary(m)$coefficients
  class(cf) <- c("coeftest", "matrix")
  cf
}
ct_ae <- fe_coeftest(m_ae)
ct_pc <- fe_coeftest(m_pc)
ct_lw <- fe_coeftest(m_lw)
ct_rw <- fe_coeftest(m_rw)

# Keep only topic fixed effects significant at p < 0.01 in at least one dimension
all_topic_coefs <- grep("^top_topic", rownames(ct_ae), value = TRUE)
p_mat <- cbind(ct_ae[all_topic_coefs, "Pr(>|z|)"], ct_pc[all_topic_coefs, "Pr(>|z|)"],
               ct_lw[all_topic_coefs, "Pr(>|z|)"], ct_rw[all_topic_coefs, "Pr(>|z|)"])
sig_topics <- all_topic_coefs[apply(p_mat, 1, min, na.rm = TRUE) < 0.01]

sig_topic_labels <- vapply(sig_topics, function(tc) {
  lbl <- topic_label_map[tc]
  if (is.na(lbl)) tc else unname(lbl)
}, character(1))

party_coefs <- grep("^Party", rownames(ct_ae), value = TRUE)
coef_names <- c(
  "(Intercept)"     = "(Intercept)",
  "Date trend (SD)" = "date_num",
  setNames(party_coefs, sub("^Party", "", party_coefs)),
  setNames(sig_topics, sig_topic_labels)
)

# ICC = σ²_speaker / (σ²_speaker + π²/3); π²/3 is the logistic level-1 variance
icc_fmt <- function(m) {
  v <- as.numeric(VarCorr(m)$Name)
  sprintf("%.3f", v / (v + pi^2 / 3))
}

suppressWarnings(huxreg(
  "Anti-Elitism"    = ct_ae,
  "People-Centrism" = ct_pc,
  "Left-Wing"       = ct_lw,
  "Right-Wing"      = ct_rw,
  coefs      = coef_names,
  statistics = character(0),
  note = paste(
    "Mixed-effects logistic regression coefficients (log-odds).",
    "Reference party: SPD.",
    "Reference topic: Agriculture.",
    "Random intercept by speaker; model-based standard errors in parentheses.",
    "Date trend standardised (coefficient = change per SD of time).",
    "Only topic fixed effects significant at p < 0.01 in at least one dimension are shown.",
    "Only sentences that have been assigned a substantive topic have been included: ",
    paste0("N = ", nobs(m_ae), " / ", nobs(m_pc), " / ", nobs(m_lw), " / ", nobs(m_rw),
           " sentences (AE / PC / LW / RW)."),
    paste0("ICC (speaker) = ", icc_fmt(m_ae), " / ", icc_fmt(m_pc), " / ",
           icc_fmt(m_lw), " / ", icc_fmt(m_rw), " (AE / PC / LW / RW)."),
    "* p < 0.1, ** p < 0.05, *** p < 0.01."
  )
)) |>
  set_caption("Mixed-Effects Logistic Regression: Predictors of Populist Rhetoric in Bundestag Speeches (20th Period)") |>
  set_label("tbl-regression") |>
  set_tabular_environment("tabular") |>
  set_font_size(8) |>
  set_width(1)

```

The intra class correlation for the model is 0.147 for anti-elitism, 0.094 for people-centrism, 0.18 for left-wing host-ideology, and 0.207 for right-wing host-ideology, meaning after controlling for all other variables in the model, 14.7% / 9.4% / 18% / 20.7% of the difference in the outcome remain due to the variance in the speakers.

In the following, only statistically significant (for p < 0.01) coefficients will be mentioned if not explicitly stated.

Over time, the log-odds for anti-elitism classification increase by 0.143 per time-step and by 0.082 for right-wing host-ideology.

Compared to the reference party SPD, the AfD shows the largest increase in the dimenstions anti-elitism and interestingly both left- and right-wing host-ideology. Being a member of the AfD increases the log-odds of a sentence being classified as anti-elitism by 1.106, as left-wing by 0.704 and as right-wing by 3.27. A similar pattern persists for MPs without a party background, which might be due to many of them stemming from the AfD or having been expelled from moderate parties.

Right-wing populism emphasize anti-elitism while far left populism focuses on people-centrism [@mudde_populism_2017; @erhard_popbert_2025]. If anti-elitism performs better on social media than people-centrism, the greater success of right-wing populism on social media becomes obvious [@mutascu_online_2025].

Similarily for MPs of the CDU/CSU as opposition parties, the log-odds increased by 0.865 to classify a sentence as anti-elitist and increased by 1.472 to classify a sentence as right-wing compared to being a member of the SPD. Different than the AfD, the CDU/CSU employed less people-centrism. The log-odds decreased by 0.332. The coefficients for Die Linke and BSW were only significant and positive for left- (Die Linke) and right-wing (Die Linke, BSW) host ideology.

While these inferential statistics are not enough to fully support the connection to social media, it does hint towards different use of populist rhetoric possibly dependent on social media strategies.
The impact of time and party affiliation is in line with right-wing populism being more successfull on social media as anti-elitism is more heavily used by right-wing parties.

The differences in coefficients for the governing parties Grüne, and FDP compared to the leading government party SPD naturally were smaller and generally had lower levels of significance indicating the reduced necessity of critizising your own coalition.

The coefficients of the topics are to be interpreted in comparison to the reference topic agriculture and controlled for party affiliation and time.

The predominance of party affairs already demonstrated in the descriptive analysis persists. Talking about other parties increases the log-odds of a sentence being classified as populist across all dimensions (0.559\*\*\*, 0.281\*\*\*, 0.636\*\*\*, 0.460\*\*). It should be mentioned that party affairs might draw significance from other topics as critizising another party often happens in the same sentence as mentioning a substantive topic.

The topic government increases the log-odds of anti-elitist classification by 0.174 and decreases in the log-odds of people centrism by 5.89 compared to agriculture suggesting that when addressing the government, critique is more focused on than people-centrism, which is expected. Similarily left-wing host-ideology has a positive, right-wing a negative effect indicating that talk about the government is embedded in left-wing host-ideology rather than right wing, at least compared to speeches about agriculture.

The emotionally charged topics security and migration show an increased level of populism. Coefficients for those are positive for people-centrism. This suggests that in these contexts, the people are put in the foreground rather than an alleged mismanagement of the government.

The likewise emotionally charged topic Environment and climate shows increasing coefficients in anti-elitism, but decreasing in people-centrism suggesting that the type of populist rhetoric used in emotionally charged topics might be dependent on the topic.

Social policy exhibits the strongest coefficient among topic effects. Compared to agriculture, when a sentence is about social policy, the log-odds of it being classified as right-wing populist decrease by 1.433 indicating that social policy is not often picked up by right-wing populists. An extension of this pattern can be observed for talk about Ukraine. Here, the coefficients of both left- and right-wing host-ideology are significant and negative.

Some level of support is found for all hypotheses, with limitations remaining in the strong fragmentation of coefficients. For hypothesis 2 and 3 a future iteration of the model could exclude party affairs and include interaction terms of parties and topics to control for effects that occur only for certain comnbinations of topics and parties.

# Conclusion

This work explored language in the German parliament and developed an argument for social media as a driver of populist rhetoric. Findings of previous literature have been supported. Populist rhetoric in the German parliament increased after the relevance of social media as political information medium increased aswell. Anti-elitism is the most dominant dimension of populist rhetoric, however, the intensity differs heavily across parties. While party affiliation seems to be the main driver for populist rhetoric, the topic of a speech can help predict in which host-ideology populism is employed. Emotionally charged topics seem to have higher levels of populism, although more in-depth analysis is needed for substantive claims.


## Limitations and future research

There are several limitations to the conclusions reached in this paper. The most obvious and important one is that no sufficient inferential analysis has been conducted to test for a relation between the relevance of social media as political information medium and the amount of populist rhetoric in parliamentary speeches.

To test if the rhetoric truly changes due to social media, data on social media posts needs to be linked directly to speeches of the Bundestag. Future research should consider using LLMs to identify populist rhetoric because they perform better when it comes to identifying nuanced stances in different contexts. Qualitative methods, especially interviews of members of parliament would also be a valuable extension.

The level of populism is evaluated by PopBERT on sentence instead of speech level. This poses two possible problems:

First, the available data from the DPI API comes on speech level. Splitting the speeches into sentences may wrongly split sentences at spots where they should not be split and vice versa. Parliamentary speeches often are no cleanly written texts. During a speech there might be interventions or slips of the tongue that lead to difficulties in determining the end and beginning of sentences.

Second, classifying on speech level would prevent missing any context information. Speech-level analysis focuses on the classifying the whole text, whereas sentence-level analysis is normally employed when concentrating on short messages [@aliyu_comparative_2024]. Two sentences on their own might not be enough to identify populist rhetoric, but when they are taken together, they might prove to be populist.

However, PopBERT is limited to 512 tokens per classification, which would require chunking speeches to avoid truncating them. This would keep more context information but the chunk might split mid-sentence which would cause again loss of information. Therefore, this work relied on classifying at sentence level, which is closest to the method used by @erhard_popbert_2025. Future research should employ a more resource intesive approach to enable classifying on speech instead of sentence level.


# References

::: {#refs}
:::

# Appendix

## Appendix A: Use of Generative AI

Every output of AI tools has been thouroughly checked before being implemented. No text passages (except for plot labels) generated by AI have been included.

```{r}
#| label: tbl-ai-tools
#| tbl-cap: "Use of generative AI tools in this research paper"
ai_appendix_table <- data.frame(
  `AI Tool`    = c("Claude Code","Claude Code", "Claude Sonnet 4.6", "Github Copilot"),
  Explanation  = c("Iterative conversation to double checked entire codebase to ensure data quality and adequate methodology.","Quality check of regression coefficient and overall feedback on the paper", "Creation of pipeline to pull Speech data from the DPI API", "used to autocomplete some sections of the code during topic analysis. Came up with parts of the keyword selection."),
  `Prompt used` = c("For full conversation see prompts_log.md. Example prompts: 'check the code for my topic modelling and tell me if and what I should do differently. do not make any changes at this point.', 'Now check the whole thing again. Is the document level topic modelling correctly implemented? Will the pipeline work without errors or bugs? Are there methodological concerns you identify? Do not change anything, but tell me about it first.'\n", "for full conversation see prompts_log.md. Example prompts: 'think about if it makes sense to remove the other_1 and other_2 topics from the regression model. they seem to have a negative effect on populist rhetoric at the moment, but i cannot meaningfully interpret them as I do not know what is talked about here.', ", "'Create a pipeline that uses following api https://dip.bundestag.de/documents/informationsblatt_zur_dip_api.pdf to get all speeches from the beginning of the 20th period of the German Bundestag. If possible, include metadata for the date, speaker and party. Do this using either R or python.'\n", "None, code autocompletion"),
  check.names  = FALSE
)

knitr::kable(ai_appendix_table, booktabs = TRUE, longtable = TRUE) |>
  kableExtra::kable_styling(full_width = TRUE) |>
  kableExtra::column_spec(1, width = "2.5cm") |>
  kableExtra::column_spec(2, width = "5.5cm") |>
  kableExtra::column_spec(3, width = "6cm")

```

## Appendix B: PopBERT performance

```{r}
#| label: tbl-popbert-performance
#| tbl-cap: "Performance of the PopBERT model on the test set [@erhard_popbert_2025]."

popbert_perf <- data.frame(
  "Dimension"  = c("Anti-elitism", "People-centrism", "Left-wing ideology",
                   "Right-wing ideology", "Micro avg", "Macro avg"),
  "Precision"  = c(0.81, 0.70, 0.69, 0.68, 0.75, 0.72),
  "Recall"     = c(0.88, 0.73, 0.77, 0.66, 0.80, 0.76),
  "F1"         = c(0.84, 0.71, 0.73, 0.67, 0.77, 0.74)
)

popbert_perf |>
  kable(
    booktabs  = TRUE,
    longtable = TRUE,
    align     = c("l", "c", "c", "c"),
    escape    = FALSE,
    digits    = 2
  ) |>
  kable_styling(
    font_size  = 9,
    full_width = FALSE
  ) |>
  row_spec(0, bold = TRUE) |>
  row_spec(4, hline_after = TRUE)
```

```{r}
#| label: fig-modelfit
#| fig-cap: "Log-likelihood over keyATM iterations, indicating model convergence."
plot_modelfit(keyatm_model)
```

## Appendix C: keyness analysis

```{r}
#| label: fig-keyness SPD
#| fig-height: 20
#| fig-width: 18

x <- keyness_by_party[["SPD"]]
textplot_keyness(x)

```

```{r}
#| label: fig-keyness Linke
#| fig-height: 20
#| fig-width: 18

x <- keyness_by_party[["Die Linke"]]
textplot_keyness(x)

```

```{r}
#| label: fig-keyness CDU/CSU
#| fig-height: 20
#| fig-width: 18

x <- keyness_by_party[["CDU/CSU"]]
textplot_keyness(x)

```

```{r}
#| label: fig-keyness B90/Grüne
#| fig-height: 20
#| fig-width: 18

x <- keyness_by_party[["B90/Die Grünen"]]
textplot_keyness(x)
names(keyness_by_party)
 
```

```{r}
#| label: fig-keyness AfD
#| fig-height: 20
#| fig-width: 18

x <- keyness_by_party[["AfD"]]
textplot_keyness(x)

```

```{r}
#| label: fig-keyness BSW
#| fig-height: 20
#| fig-width: 18

x <- keyness_by_party[["BSW"]]
textplot_keyness(x)

```

```{r}
#| label: fig-keyness fraktionslos
#| fig-height: 20
#| fig-width: 18

x <- keyness_by_party[["fraktionslos"]]
textplot_keyness(x)

```


